---
output: 
html_document: 
keep_md: yes
---
# **Reproducible Research Project 1**
        
        
### 1-Code for reading in the dataset and/or processing the data
        
```{r}
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(lubridate))
        
activity <- read.csv("repdata_data_activity/activity.csv")
str(activity); summary(activity); colSums(is.na(activity))
```
        
### 2-Histogram of the total number of steps taken each day
        
```{r}
steps_per_day <- activity %>%
        group_by(date) %>%
        summarize(total_steps = sum(steps ,na.rm = T))
```
        
```{r fig.height= 4 ,fig.width=8}
ggplot(steps_per_day ,aes(total_steps))+
        geom_histogram(bins = 30 ,fill="purple" ,alpha=0.4)+
        ggtitle("Total steps per day")+ xlab("Toal steps")+
        theme_minimal()
```
        
        
## 3-Mean and median number of steps taken each day
        
```{r}
mean_day_steps <-mean(steps_per_day$total_steps ,na.rm = TRUE)
median_day_steps <- median(steps_per_day$total_steps ,na.rm = TRUE)
        
mean_day_steps ; median_day_steps
```
        
## 4-Time series plot of the average number of steps taken
        
```{r fig.height= 4 ,fig.width=8}
average_steps <- activity %>%
        group_by(interval)%>%
        summarise(mean_stps =mean(steps ,na.rm = TRUE))
        
ggplot(average_steps)+
        geom_line(aes(x=interval ,y=mean_stps) ,color="purple")+
        ggtitle("Time Series Plot of Average Steps")+
        ylab("Average Steps") + xlab("Intervals")+theme_minimal() 
```
        
## 5-The 5-minute interval that, on average, contains the maximum number of steps
        
```{r}
average_steps[which.max(average_steps$mean_stps),]
```
        
## 6- Code to describe and show a strategy for imputing missing data
        
```{r}
mean_new_steps <- mean(activity$steps ,na.rm = TRUE)
steps_without_missing <- ifelse(is.na(activity$steps), mean_new_steps ,activity$steps)
sum(is.na(steps_without_missing))
new.data <- cbind(activity[,c("date" ,"interval")],steps_without_missing)
head(new.data)
```
        
        
## 7-Histogram of the total number of steps taken each day after missing values are imputed
        
```{r}
complete_steps_per_day <- new.data %>%
        group_by(date) %>%
        summarise(stps = sum(steps_without_missing ))
```

```{r fig.height= 4 ,fig.width=8}
ggplot(complete_steps_per_day)+
        geom_histogram(aes(x=stps),binwidth = 1000 ,fill="purple" ,alpha=0.5)+
        ggtitle("Total number of steps per day after missing values are imputed")+
        xlab("Steps")+theme_minimal()
```
        
## 8-Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
        
```{r}
is_weekday <-function(date){
        ifelse(wday(date)%in%c(1,7),"weekend","weekday")
                
}
new.data <- mutate(new.data,date=ymd(date)) %>% mutate(day=sapply(date,is_weekday))
head(new.data)
```
        
```{r}
averages <- new.data %>%
        group_by(interval ,day) %>%
        summarise(avr_steps = mean(steps_without_missing))
```     

```{r fig.height= 4 ,fig.width=8}
ggplot(averages ,aes(interval , avr_steps))+
        geom_line(color="purple")+
        facet_wrap(day~.)+
        xlab("Intervals")+ ylab("Steps")+
        ggtitle("Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends")+
        theme_minimal()
```
